---
title: "Example package results spec"
author: "James P. Gilbert"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{ExamplePackage}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r,echo=FALSE}
specification <- data.frame(
  namespace = "sfe",
  namespacePrefix = "sfe_",
  tableName = c("covariate_definition", "covariate_definition", "covariate_result", "covariate_result", "covariate_result"),
  columnName = c("covariate_id", "covariate_name", "cohort_definition_id", "covariate_id", "covariate_mean"),
  dataType = c("int", "varchar", "int", "bigint", "numeric"),
  primaryKey = c("Yes", "No", "Yes", "Yes", "No"),
  nullable = c("No", "No", "No", "No", "No"),
  isRequired = c("Yes", "Yes", "Yes", "Yes", "Yes"),
  optional = c("No", "No", "No", "No", "No")
)
```

# Introduction
This guide intends to serve as an example of using RMM to build and maintain a package that produces results in an
end to end manner.
The aspects this package will cover are as follows:

* Creating a basic package results specification
* Using this specification to create a database schema and instantiating it in an SQLite database
* Creating a database migration for this project

# Setup R for data export

## Project setup

First we will create an R package called `SimpleFeatureExtractor` a toy example that pulls a set of aggregate features
for specified cohorts from an OMOP CDM and exports the result set to a relational database for further analysis.

In this example we will export two tables:

| namespace | table_name           | column_name           | data_type | primary_key | nullable |
|-----------|----------------------|-----------------------|-----------|-------------|----------|
| sfe       | covariate_definition | covariate_id          | int       | Yes         | No       |
| sfe       | covariate_definition | covariate_name        | varchar   | No          | No       |
| sfe       | covariate_result     | cohort_definition_id  | int       | Yes         | No       |
| sfe       | covariate_result     | covariate_id          | bigint    | Yes         | No       |
| sfe       | covariate_result     | covariate_mean        | numeric   | No          | No       |

Results exported from this package are covariate prevalences related to cohorts, given a covariate_id, these are
related to names in a second table.

The recommended format is a namespaced YAML file saved to the `inst/settings/` folder of your R package.
This file should be called `resultsDataModelSpecification.yaml`:

```yaml
version: "1.0"

namespace:
  sfe:
    prefix: sfe_
    description: "Simple Feature Extractor results"
    tables:
      covariate_definition:
        columns:
          - name: covariate_id
            type: int
            primary_key: true
            description: "The unique covariate identifier"
          - name: covariate_name
            type: varchar
            description: "Human-readable covariate name"
      covariate_result:
        columns:
          - name: cohort_definition_id
            type: int
            primary_key: true
            references: sfe_covariate_definition.covariate_id
          - name: covariate_id
            type: bigint
            primary_key: true
            references: sfe_covariate_definition.covariate_id
          - name: covariate_mean
            type: numeric
```

Each HADES package owns its own YAML specification. There is no central results model repository —
packages ship their schema definitions independently. Combining multiple specs into a unified
schema is handled by higher-level orchestrators (e.g. Strategus).

CSV format is still supported but deprecated; use `csvToYaml()` to migrate existing CSV files.

The package should create results csv files that correspond to these fields in terms of type and name.



## Creating a results database schema
First we should load our specification
```{r,eval=FALSE}
specification <- ResultModelManager::loadResultsDataModelSpecifications(
  system.file("settings", "resultsDataModelSpecification.yaml", package = "SimpleFeatureExtractor")
)
```

We can then create our schema from this sql:

```{r}
sql <- ResultModelManager::generateSqlSchema(schemaDefinition = specification)
```
Viewing the sql we can see that we should add a `database_schema` parameter when executing the sql and `table_prefix` if
we need it.
```{r}
writeLines(sql)
```
We can then easily use this to create a schema using a `QueryNamespace`:
```{r}
connectionDetails <- DatabaseConnector::createConnectionDetails(
  dbms = "sqlite",
  server = tempfile()
)
qns <- ResultModelManager::createQueryNamespace(
  connectionDetails = connectionDetails,
  tableSpecification = specification,
  tablePrefix = "my_study_",
  database_schema = "main"
)

# The schema DDL references namespaced table names (e.g. @sfe_covariate_definition)
qns$executeSql(sql)
```
Alternatively, we can just use `DatabaseConnector` functions directly.
```{r,eval=FALSE}
connection <- DatabaseConnector::connect(connectionDetails)
DatabaseConnector::renderTranslateExecuteSql(connection,
  sql,
  table_prefix = "my_study_",
  database_schema = "main"
)
```
## Uploading results

Now we have a schema we can upload results to it using the functionality exposed in this package.
Using the above example, a results folder should have the following files:

`results/:`

| File Name                 | Description                 |
|---------------------------|-----------------------------|
| covariate_definition.csv  | Covariate Definition File    |
| covariate_result.csv      | Covariate Result File        |


We can now use the results spec to upload these files (and validate that they conform to the specification):
```{r,eval=FALSE}
ResultModelManager::uploadResults(connectionDetails,
  schema = "main",
  resultsFolder = "results",
  tablePrefix = "my_study_",
  specifications = specification
)
```

With the results uploaded we can now write queries inside the namespace using the namespaced table references:

```{r}
qns$queryDb("SELECT * FROM @database_schema.@sfe_covariate_definition")
```

```{r}
qns$queryDb("SELECT * FROM @database_schema.@sfe_covariate_result WHERE cohort_definition_id = @cohort_id",
  cohort_id = 5
)
```
