---
title: "Example package results spec"
author: "James P. Gilbert"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{ExamplePackage}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r,echo=FALSE}
specification <- data.frame(
  tableName = c("covariate_definition", "covariate_definition", "covariate_result", "covariate_result", "covariate_result"),
  columnName = c("covariate_id", "covariate_name", "cohort_definition_id", "covariate_id", "covariate_mean"),
  dataType = c("int", "varchar", "int", "bigint", "numeric"),
  isRequired = c("Yes", "Yes", "Yes", "Yes", "Yes"),
  primaryKey = c("Yes", "No", "Yes", "Yes", "No"),
  optional = c("No", "No", "No", "No", "No"),
  emptyIsNa = c("No", "No", "No", "No", "No")
)
```

# Introduction
This guide intends to server as an example of using RMM to build and maintain a package that produces results in an
end to end manner.
The aspects this package will cover are as follows:

* Creating a basic package results specification
* Using this specification to create a database schema and instantiating it in an SQLite database
* Creating a database migration for this project

# Setup R for data export

## Project setup

First we will create an R package called `SimpleFeatureExtractor` a toy example that pulls a set of aggregate features
for specified cohorts from an OMOP CDM and exports the result set to a relational database for further analysis.

In this example we will export a single csv file that contains the following:

| table_name           | column_name           | data_type | is_required | primary_key | optional | empty_is_na |
|----------------------|-----------------------|-----------|-------------|-------------|----------|-------------|
| covariate_definition | covariate_id          | int       | Yes         | Yes         | No       | No          |
| covariate_definition | covariate_name        | varchar   | Yes         | No          | No       | No          |
| covariate_result     | cohort_definition_id  | int       | Yes         | Yes         | No       | No          |
| covariate_result     | covariate_id          | bigint    | Yes         | Yes         | No       | No          |
| covariate_result     | covariate_mean        | numeric   | Yes         | No          | No       | No          |


Results exported from this package are covariate prevalances related to a cohorts and given a covariate_id, these are
related to names in a second table.
This table should be saved to the `inst` folder of your R pacakge.
Preferably, this file should be called `resultsDataModelSpecification.csv`

The package should create results csv files that correspond to these fields in terms of type and name.



## Creating a results database schema
First we should load our specification
```{r,eval=FALSE}
specification <- ResultModelManager::loadResultsDataModelSpecifications("resultsDataModelSpecification")
```

We can then create our schema from this sql:

```{r}
sql <- ResultModelManager::generateSqlSchema(schemaDefinition = specification)
```
Viewing the sql we can see that we should add a `database_schema` parameter when executing the sql and `table_prefix` if
we need it.
```{r}
writeLines(sql)
```
We can then easily use this to create a schema using a `QueryNamespace`:
```{r}
connectionDetails <- DatabaseConnector::createConnectionDetails(dbms = "sqlite",
                                                                server = tempfile())
qns <- ResultModelManager::createQueryNamespace(connectionDetails = connectionDetails,
                                                tableSpecification = specification,
                                                tablePrefix = "my_study_",
                                                database_schema = "main")

# note - the table prefix and schema parameters are not neeeded when we do this
qns$executeSql(sql)
```
Alternatively, we can just use `DatabaseConnector` functions directly.
```{r,eval=FALSE}


connection <- DatabaseConnector::connect(connectionDetails)
DatabaseConnector::renderTranslateExecuteSql(connection,
                                             sql,
                                             table_prefix = "my_study_",
                                             database_schema = "main")


```
## Uploading results

Now we have a schema we can upload results to it using the functionality exposed in this package.
Using the above example, a results folder should have the following files:

`results/:`

| File Name                 | Description                 |
|---------------------------|-----------------------------|
| covariate_definition.csv  | Covariate Definition File    |
| covariate_result.csv      | Covariate Result File        |


We can now use the results spec to upload these files (and validate that they conform to the specification):
```{r,eval=FALSE}
ResultModelManager::uploadResults(connectionDetails,
                                  schema = "main",
                                  resultsFolder = "results",
                                  tablePrefix = "my_study_",
                                  specifications = specification)
```

With the results uploaded we can now write queries inside the namespace:

```{r}
qns$queryDb("SELECT * FROM @database_schema.@covariate_definition")

```

```{r}
qns$queryDb("SELECT * FROM @database_schema.@covariate_result WHERE cohort_definition_id = @cohort_id",
            cohort_id = 5)

```