---
title: "Upload Functionality"
author: "James P. Gilbert"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{Upload Functionality}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction
This vignette describes the functionality for uploading results to a pre-created database schema.
In the examples here, we assume the use of sqlite for simplicity.
However, in principle any platform supported by the `DatabaseConnector` and `SqlRender` packages should work.

# Creating a schema definition file
Every analytics package that creates data output should contain a **YAML specification** that defines
its results data model. The format is a namespaced YAML schema; CSV support is deprecated
(use `csvToYaml()` to migrate existing CSV files).

## Schema ownership philosophy

Each HADES package ships its own `resultsDataModelSpecification.yaml` in its `inst/settings/` directory.
A package's YAML defines **only the tables produced by that package**.

This avoids the circular dependency problem of a central results model repository:
packages do not need to know about each other's schemas. Combining specifications from multiple
packages into a unified HADES results schema is the responsibility of a higher-level orchestrator
such as Strategus.

## YAML format (recommended)

Schema definitions must include the following information per column:

- **column name** and **data type** (e.g., `varchar`, `bigint`, `int`, `float`, `date`, `varchar(255)`)
- **primary key** indicator
- **nullable** flag (whether the column may contain NULL values)

Additional optional fields:

- `min_cell_count` — marks columns requiring minimum cell count enforcement for privacy
- `empty_is_na` — whether empty strings should be treated as NA (default: true)
- `description` — human-readable description of the column
- `references` — foreign key reference in `table.column` format

A fully-featured namespaced YAML specification looks like this:

```yaml
version: "1.0"

namespace:
  my_analysis:
    prefix: my_analysis_        # database table prefix (defaults to {namespace}_)
    description: "Results from the MyAnalysis package"
    tables:
      table_1:
        description: "Primary analysis results"
        columns:
          - name: database_id
            type: varchar
            primary_key: true
            description: "The database identifier"
          - name: analysis_id
            type: bigint
            primary_key: true
          - name: analysis_name
            type: varchar
          - name: domain_id
            type: varchar(20)
            nullable: true
          - name: start_day
            type: float
            nullable: true
          - name: end_day
            type: float
            nullable: true
          - name: is_binary
            type: varchar(1)
          - name: missing_means_zero
            type: varchar(1)
            nullable: true
      table_2:
        columns:
          - name: database_id
            type: varchar
            primary_key: true
            references: table_1.database_id
          - name: analysis2_id
            type: bigint
            primary_key: true
          - name: concept_id
            type: int
          - name: logic_description
            type: varchar
            nullable: true
          - name: valid_start_date
            type: Date
          - name: concept_name
            type: varchar(255)
          - name: p_10_value
            type: float
```

The namespace name (and `prefix`) determine the database table names.
A table `table_1` in namespace `my_analysis` with prefix `my_analysis_` will be created as
`my_analysis_table_1`. The `prefix` field is optional and defaults to `{namespace}_`.

## Platform-specific configuration

Platform-specific features such as table partitioning and indexes can be configured
per DBMS platform in the same YAML file:

```yaml
platforms:
  postgresql:
    namespace:
      my_analysis:
        tables:
          table_1:
            partition_by: RANGE (analysis_id)
            indexes:
              - columns: [database_id, analysis_id]
                unique: true
  sql_server:
    namespace:
      my_analysis:
        tables:
          table_1:
            indexes:
              - columns: [database_id]
  duckdb: {}
  sqlite: {}
```

Features not supported by a platform (e.g., partitioning in duckdb) will emit a warning.

## CSV format (deprecated)

```{r eval=FALSE}
# File inst/settings/resulsDataModelSpecifications.csv
table_name,column_name,data_type,is_required,primary_key
table_1,database_id,varchar,Yes,Yes
table_1,analysis_id,bigint,Yes,Yes
...
```

CSV-based specifications are deprecated. Use `csvToYaml()` to migrate existing CSV files
to the YAML format.

## Loading specifications

To load a specification file:

```{r eval = FALSE}
# YAML (recommended) - returns a data.frame
spec <- ResultModelManager::loadResultsDataModelSpecifications("path/to/spec.yaml")

# YAML with platform config - returns list(specification, platforms)
result <- ResultModelManager::loadResultsDataModelFromYaml("path/to/spec.yaml")
spec <- result$specification      # data.frame with namespace, tableName, ...
platforms <- result$platforms     # platform-specific config

# CSV (deprecated) - emits a deprecation warning
spec <- ResultModelManager::loadResultsDataModelSpecifications("path/to/spec.csv")
```

For loading inside an R package via `system.file`:

```{r eval = FALSE}
#' Get Results Data Model Specifications
getResultsDataModelSpec <- function() {
  specPath <- system.file("settings", "resultsDataModelSpecification.yaml",
                          package = utils::packageName())
  spec <- ResultModelManager::loadResultsDataModelSpecifications(specPath)
  return(spec)
}
```

# Creating a schema
This section describes how to use `RMM` to create a schema from a specifications file.

```{r eval=FALSE}
connectionDetails <- DatabaseConnector::createConnectionDetails("sqlite", server = "MySqliteDb.sqlite")
connection <- DatabaseConnector::connect(connectionDetails)
sql <- ResultModelManager::generateSqlSchema(schemaDefinition = getResultsDataModelSpec())
DatabaseConnector::renderTranslateExecuteSql(connection, sql, database_schema = "main", table_prefix = "pre_")
DatabaseConnector::disconnect(connection)
```

Note that the SQL generated by `generateSqlSchema` contains the required parameter `database_schema` and the optional
parameter `table_prefix`.

If the specification was loaded from a YAML file (not a data.frame), you can pass a `platform` parameter
for platform-specific DDL (partitioning, indexes):

```{r eval=FALSE}
sql <- ResultModelManager::generateSqlSchema(
  csvFilepath = "path/to/spec.yaml",
  platform = "postgresql"
)
```

# Uploading results
This section shows how to upload results conforming to the above specified schema.
It is assumed that a zip file has been created with csv files corresponding to the table names provided in the
`resultsDataModelSpecification` file.

```{r eval=FALSE}
ResultModelManager::unzipResults(zipFile = "MyResultsZip.zip", resultsFolder = "extraction_folder")
ResultModelManager::uploadResults(connectionDetails,
  schema = "main",
  resultsFolder = "extraction_folder",
  tablePrefix = "pre_",
  purgeSiteDataBeforeUploading = FALSE,
  specifications = getResultsDataModelSpec()
)
```
This will extract the zip file and upload results using `DatabaseConnector`.
For most platforms some form of bulk loading is supported, the setup for this is contained in the documentation for
`DatabaseConnector`, it is strongly advised that you set up and use this functionality for any platforms where you will
store large data sets.
However, this is not necessary for the sqlite example demonstrated here.
