---
title: "Upload Functionality"
author: "James P. Gilbert"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{Upload Functionality}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction
This vignette describes the functionality for uploading results to a pre-created database schema.
In the examples here, we assume the use of sqlite for simplicity.
However, in principle any platform supported by the `DatabaseConnector` and `SqlRender` packages should work.

# Creating a schema definition file
Every analytics package that creates data output should contain a YAML or CSV specification that defines
its results data model. This package uses a **namespaced YAML format** by default; CSV support is deprecated.

Schema definitions must include the following information per column:

- **table name** (within a namespace)
- **column name**
- **data type** (e.g., `varchar`, `bigint`, `int`, `float`, `date`, `varchar(255)`)
- **primary key** indicator

In addition, optional fields such as `optional`, `min_cell_count`, `empty_is_na`, and `description`
may be included.

## YAML format (recommended)

A namespaced YAML specification looks like this:

```yaml
version: "1.0"

namespace:
  my_analysis:
    tables:
      table_1:
        columns:
          - name: database_id
            type: varchar
            primary_key: true
          - name: analysis_id
            type: bigint
            primary_key: true
          - name: analysis_name
            type: varchar
          - name: domain_id
            type: varchar(20)
            optional: true
          - name: start_day
            type: float
            optional: true
          - name: end_day
            type: float
            optional: true
          - name: is_binary
            type: varchar(1)
          - name: missing_means_zero
            type: varchar(1)
            optional: true
      table_2:
        columns:
          - name: database_id
            type: varchar
            primary_key: true
          - name: analysis2_id
            type: bigint
            primary_key: true
          - name: concept_id
            type: int
          - name: logic_description
            type: varchar
            optional: true
          - name: valid_start_date
            type: Date
          - name: concept_name
            type: varchar(255)
          - name: p_10_value
            type: float
```

The namespace name (e.g. `my_analysis`) becomes the table prefix in the database.
A table named `table_1` in namespace `my_analysis` will be created as `my_analysis_table_1`.

Platform-specific features such as table partitioning and indexes can be configured
per DBMS platform. See `?loadResultsDataModelFromYaml` for details.

## CSV format (deprecated)

```{r eval=FALSE}
# File inst/settings/resulsDataModelSpecifications.csv
table_name,column_name,data_type,is_required,primary_key
table_1,database_id,varchar,Yes,Yes
table_1,analysis_id,bigint,Yes,Yes
...
```

CSV-based specifications are deprecated. Use `csvToYaml()` to migrate existing CSV files
to the YAML format.

## Loading specifications

To load a specification file:

```{r eval = FALSE}
# YAML (recommended)
spec <- ResultModelManager::loadResultsDataModelSpecifications("path/to/spec.yaml")

# The YAML specification is loaded via
result <- ResultModelManager::loadResultsDataModelFromYaml("path/to/spec.yaml")
spec <- result$specification  # data.frame
platforms <- result$platforms  # platform-specific config

# CSV (deprecated)
spec <- ResultModelManager::loadResultsDataModelSpecifications("path/to/spec.csv")
```

For loading inside an R package via `system.file`:

```{r eval = FALSE}
#' Get Results Data Model Specifications
getResultsDataModelSpec <- function() {
  specPath <- system.file("settings", "resultsDataModelSpecification.yaml",
                          package = utils::packageName())
  spec <- ResultModelManager::loadResultsDataModelSpecifications(specPath)
  return(spec)
}
```

# Creating a schema
This section describes how to use `RMM` to create a schema from a specifications file.

```{r eval=FALSE}
connectionDetails <- DatabaseConnector::createConnectionDetails("sqlite", server = "MySqliteDb.sqlite")
connection <- DatabaseConnector::connect(connectionDetails)
sql <- ResultModelManager::generateSqlSchema(schemaDefinition = getResultsDataModelSpec())
DatabaseConnector::renderTranslateExecuteSql(connection, sql, database_schema = "main", table_prefix = "pre_")
DatabaseConnector::disconnect(connection)
```
Note that the SQL generated by `generateSqlSchema` contains the required parameter `database_schema` and the optional
parameter `table_prefix`.

# Uploading results
This section shows how to upload results conforming to the above specified schema.
It is assumed that a zip file has been created with csv files corresponding to the table names provided in the
`resultsDataModelSpecification` file.

```{r eval=FALSE}
ResultModelManager::unzipResults(zipFile = "MyResultsZip.zip", resultsFolder = "extraction_folder")
ResultModelManager::uploadResults(connectionDetails,
  schema = "main",
  resultsFolder = "extraction_folder",
  tablePrefix = "pre_",
  purgeSiteDataBeforeUploading = FALSE,
  specifications = getResultsDataModelSpec()
)
```
This will extract the zip file and upload results using `DatabaseConnector`.
For most platforms some form of bulk loading is supported, the setup for this is contained in the documentation for
`DatabaseConnector`, it is strongly advised that you set up and use this functionality for any platforms where you will
store large data sets.
However, this is not necessary for the sqlite example demonstrated here.
