---
title: "Using Query Namespaces"
author: "James P. Gilbert"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{UsingQueryNamespaces"}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Purpose
The `QueryNamespace` class is designed to be a convenient way to write (and re-write) SQL queries for packages with
defined result model specifications.
The convenience is that only passed parameters must be set in the query - any table names or other pre-defined variables
can be set once in a result spec and reused if table prefixes are applied to the tables.
The intention is to save time and limit bugs.
This also builds on the `SqlRender`/`DatabaseConnector` principle of "Write sql once, use anywhere" principle across
ohdsi packages.
This is not intended to replace usage of `dbplyr` style operations which are expressive and allow use of sql.
However, many find that writing SQL strings is often more convenient and portable to other programming language
than `dplyr` calls allow.

# Namespaced table specs

For packages that ship YAML specifications, the table names include a namespace prefix
(see the "Upload Functionality" vignette for the YAML schema format).
With a namespaced spec, the `QueryNamespace` registers tables as `@namespace_tableName`:

```{r}
library(ResultModelManager)

# Namespaced table spec (from a YAML file with namespace = "cg")
tableSpecification <- data.frame(
  namespace = "cg",
  namespacePrefix = "cg_",
  tableName = "cohort_definition",
  columnName = c("cohort_definition_id", "cohort_name", "json", "sql"),
  primaryKey = c("yes", "no", "no", "no"),
  dataType = c("bigint", "varchar", "varchar", "varchar")
)

connectionDetails <- DatabaseConnector::createConnectionDetails("sqlite", server = tempfile())
qns <- createQueryNamespace(
  connectionDetails = connectionDetails,
  usePooledConnection = FALSE,
  tableSpecification = tableSpecification,
  tablePrefix = "rwe_study_99_",
  snakeCaseToCamelCase = TRUE,
  database_schema = "main"
)

# Create our schema within the namespace
sql <- generateSqlSchema(schemaDefinition = tableSpecification)
# note - the table prefix and schema parameters are not needed
qns$executeSql(sql)
```
We can then query the table with sql that automatically replaces the table names:

```{r}
qns$queryDb("SELECT * FROM @database_schema.@cg_cohort_definition")
```
Note that the underlying query is already handling our `tablePrefix` for us,
so we don't need to add it. Here `@cg_cohort_definition` resolves to
`rwe_study_99_cg_cohort_definition`.

# Adding replacement variables at runtime
Variables can naturally be added at runtime, for example, in a query:
```{r}
qns$queryDb("SELECT * FROM @database_schema.@cg_cohort_definition WHERE cohort_definition_id = @id",
  id = 5
)
```
Alternatively we can persist the id in the object for use in all queries.

```{r}
qns$addReplacementVariable("database_id", "my_cdm")
```
Note that replacing the same variable will result in an error
```{r, eval = FALSE}
qns$addReplacementVariable("database_id", "my_cdm")
```

# Non-namespaced table specs (backward compat)

For backward compatibility, table specs without a `namespace` column work the same way
as before, registering tables as `@tableName`:

```{r}
# Non-namespaced spec - separate QueryNamespace
connectionDetails2 <- DatabaseConnector::createConnectionDetails("sqlite", server = tempfile())

tableSpecification2 <- data.frame(
  tableName = "database_info",
  columnName = c("database_id", "database_name"),
  primaryKey = c("yes", "no"),
  dataType = c("varchar", "varchar")
)

qns2 <- createQueryNamespace(
  connectionDetails = connectionDetails2,
  tableSpecification = tableSpecification2,
  tablePrefix = "legacy_",
  snakeCaseToCamelCase = TRUE,
  database_schema = "main"
)

sql2 <- generateSqlSchema(schemaDefinition = tableSpecification2)
qns2$executeSql(sql2)
```
Tables registered without a namespace are queried by their plain name:
```{r}
qns2$queryDb("SELECT * FROM @database_schema.@database_info")
```
